ERSPAN Support for Linux

### 2018-10-15 paper.tex
work on section 3

The ERSPAN tunnel is represented in Linux as a netdev and configured
through iproute2~\cite{iproute2}. Any packet that is placed into its send
queue will be encapsulated based on the netdev's ERSPAN configuration.  
As a result on the right-most, any other linux netdev which wants to create
a ERSPAN mirrored packet simply make a copy and forward to the ERSPAN netdev.
For example, a physical netdev can use TC~\cite{tc} to copy a packet
to the erspan tunnel~\cite{tcselftest}.

Mirrored traffic arrives at the sniffer machine needs to be able to
extract and restore the original monitored frame.  To differentiated
the three use cases, the administrator can create three ERSPAN session
ID, a configuration parameter for grouping the mirrored traffic.
For Linux user, an ERSPAN tunnel can also be used at the sniffer size.
Any packet arriving at the ERSPAN tunnel netdev's receive queue will
be decapsulated. Tools such as Wireshark~\cite{wireshark} can be used
to inspect the mirrored packet.

\section{ERSPAN Protocol Implementation}
\begin{figure}
{\scriptsize
\begin{verbatim}
<------------ outer -------------->  <---- inner ----...
+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
| Ether |  IP  |  GRE  |  ERSPAN  | Ether |  IP  | ...   
+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
\end{verbatim}
}
\vspace{-1.0em}
\caption{An example of mirrored packet with outer header containing
the GRE and ERSPAN header, followed by the innter Ethernet frame.}
\label{erspanhdr}
\vspace{-1.0em}
\end{figure}

@@

\subsection{Native v.s Metadata-Mode Tunnel}
%https://netdevconf.org/1.2/papers/borkmann.pdf
There are two tunnel type implementations in Linux kernel:
native tunnel and metadata-mode tunnel~\cite{lwtunnel}.
Native tunnel is the basic way of creating tunnels in Linux.
A tunnel netdev is created with per tunnel-specific configuration,
tied together with the netdev. For example, creating a GRE tunnel
with key and sequence number can be done by:
\texttt{ip link add dev gre123 type gretap local 1.1.1.1 remote 2.2.2.2 seq key 0xfb}.
As a result, $N$ different tunnel configurations require creating
$N$ number of netdevs.  In certain cases such as network virtualization,
this is not scalable because every host in the network creates
mutiple tunnels with different configurations to every other hosts~\cite{mvp}.

@@

We use Figure~\ref{example} as an example topology to demostrate
the three configuration ways of ERSPAN.
Assuming a network administrator wants to monitor a network consists
of 1) physical machines connected to Cisco switches, 2)
virtualized Linux machine with multiple VMs deployed and virtual switch
(openvswitch.ko) enforcing the fowarding policies, and
3) non-virtualized Linux physical machines, e.g., service nodes
in data center such as gateways.
Assuming all the monitored servers are under IP subnet of $10.1.1.*$,
and We place a traffic sniifer over another IP network of $192.168.1.*$.
The following subsection describes the configuration for each case.

\subsection{Cisco Switch}
We use Nexus 5000 switch and configure its ERSPAN tunnel
on ports 11 and 12 as below.
\begin{verbatim}
    monitor session 10 type erspan-source
      erspan-id 10
      vrf default
      destination ip 192.168.1.1
      source interface Ethernet1/11 both
      source interface Ethernet1/12 both
      no shut
    monitor erspan origin ip-address 192.168.1.2 global
\end{verbatim}

\subsection{openvswitch.ko}\label{ovs}
Open vSwitch consists of a userspace daemon, ovs-vswitchd, and a kernel
datapath implementation, openvswitch.ko.
This example shows a case where we only use the netlink API and openvswitch.ko
module, without the ovs-vswitchd daemon.
ovs-dpctl is a stand-alone tool which programs openvswitch.ko
through network api, see Joes' paper
\begin{verbatim}
ip link add dev myerspan type erspan external
ovs-dpctl add-dp mydp
ovs-dpctl add-if mydp myerspan
ovs-dpctl add-if mydp veth1 // connected to VM1
ovs-dpctl add-flow \
    "in_port(2)" \  // any packets matching in_port=2
    "set(tunnel(tun_id=20,dst=192.168.1.1,ttl=64,\
    erspan(ver=2,dir=1,hwid=0x4),flags(df|key))),3"
\end{verbatim}

>>>>

The ERSPAN tunnel is represented in Linux as a netdev and configured
through iproute2~\cite{iproute2}. Any packet that is placed into its send
queue will be encapsulated based on the netdev's ERSPAN configuration.  
As a result on the right-most, any other linux netdev which wants to create
a ERSPAN mirrored packet simply make a copy and forward to the ERSPAN netdev.
For example, a physical netdev can use linux TC~\cite{tcsubsys} with mirror
action to copy a packet to the erspan tunnel.

Mirrored traffic arrives at the sniffer machine needs to be able to
extract and restore the original monitored frame.  To differentiated
the three use cases, the administrator can create three ERSPAN session
ID, a configuration parameter for grouping the mirrored traffic.
For Linux user, an ERSPAN tunnel can also be used at the sniffer size.
Any packet arriving at the ERSPAN tunnel netdev's receive queue will
be decapsulated. Tools such as Wireshark~\cite{wireshark_erspan,wireshark_erspan2}
can be used to inspect the mirrored packet.

\section{ERSPAN Protocol Implementation}
\begin{figure}
{\scriptsize
\begin{verbatim}
<------------ outer -------------> <---- inner ---- ...
+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
| Ether |  IP  |  GRE  |  ERSPAN  | Ether |  IP  | ...   
+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
\end{verbatim}
}
\vspace{-1.0em}
\caption{An example of mirrored packet with outer header containing
the GRE and ERSPAN header, followed by the innter Ethernet frame.}
\label{erspanhdr}
\vspace{-1.0em}
\end{figure}

@@

\subsection{Native v.s Metadata-Mode Tunnel}
%https://netdevconf.org/1.2/papers/borkmann.pdf
There are two tunnel type implementations in Linux kernel:
native tunnel and metadata-mode tunnel~\cite{lwtunnel}.
Native tunnel is the basic way of creating tunnels in Linux.
A tunnel netdev is created with per tunnel-specific configuration,
tied together with the netdev. For example, creating a GRE tunnel
with key and sequence number can be done by:
\texttt{ip link add dev gre123 type gretap local 1.1.1.1 remote 2.2.2.2 seq key 0xfb}.
As a result, $N$ different tunnel configurations require creating
$N$ number of netdevs.  In certain cases such as network virtualization,
this is not scalable because every host in the network creates
mutiple tunnels with different configurations to every other hosts~\cite{nvp}.

@@

We use Figure~\ref{example} as an example topology to demostrate
the three configuration ways of ERSPAN.
Assuming a network administrator wants to monitor a network consists
of 1) physical machines connected to Cisco switches, 2)
virtualized Linux machine with multiple VMs deployed and virtual switch
(openvswitch.ko) enforcing the fowarding policies, and
3) non-virtualized Linux physical machines, e.g., service nodes
in data center such as gateways.
Assuming all the monitored servers are under IP network of $10.1.1.*$,
and We place a traffic sniifer over another IP network of $192.168.1.*$.
The following subsection describes the configuration for each case.

\subsection{Cisco Switch}
We use Nexus 5000 switch and configure its ERSPAN tunnel
with session ID 10, remote IP pointing to the traffic sniffer, $192.168.1.1$,
and local IP address as $192.168.1.2$.  As a result, both the ingress and
egress traffic on ports 11 and 12 will be mirrored to the remote sniffer.
\begin{verbatim}
    monitor session 10 type erspan-source
      erspan-id 10
      vrf default
      destination ip 192.168.1.1
      source interface Ethernet1/11 both
      source interface Ethernet1/12 both
      no shut
    monitor erspan origin ip-address 192.168.1.2 global
\end{verbatim}

\subsection{Open vSwitch Kernel Module}\label{ovs}
Open vSwitch consists of two components: a userspace daemon, called ovs-vswitchd,
and a flow cache as a kernel module, called openvswitch.ko.  While ovs-vswitchd
talks to the OpenFlow controller and programs its OpenFlow flow tables,
the openvswitch.ko keeps a cache where the subsequent flows are handled
inside the kernel space.

The openvswitch.ko provides a user-facing netlink API that models a network
bridge that connects multiple ports through a single table~\cite{ovswoovs}.
This example shows how to use the netlink API provided by openvswitch.ko
module, with the utility, ovs-dpctl, to create a ERSPAN tunnel.
\begin{verbatim}
# creating datapath named "mydp", attach veth1(port 2)
ovs-dpctl add-dp mydp
ovs-dpctl add-if mydp veth1 // connected to VM1

# creating erspan dev named "myerspan" and attach
ip link add dev myerspan type erspan external
ovs-dpctl add-if mydp myerspan

# flow entry for port 2 to erspan tunnel 
ovs-dpctl add-flow \
    "in_port(2)" \  // any packets matching in_port=2
    "set(tunnel(tun_id=20,dst=192.168.1.1,ttl=64,\
    erspan(ver=2,dir=1,hwid=0x4),flags(df|key))),3"
\end{verbatim}

%\begin{lstlisting}[caption={Example ovs-dpctl usage},language=bash,label=lst:dpctl,breaklines=true]
%\end{lstlisting}
